---
title: "Stats 485 Unit 3 Appendix"
author: "Emily MacQuarrie"
date: "4/26/2019"
output: html_document
---

#Load Libraries
```{r}
library(dplyr)
library(ggplot2)
library(lme4)
# suppress warnings
options(warn=-1)
```

# Load Data
```{r}
set.seed(485)
security_wide <- read.csv('http://dept.stat.lsa.umich.edu/~bbh/s485/data/security_wide.csv')
nschools <-nlevels(security_wide$school)
# Create dataframe for analysis
fs_items <- c(paste0('q38',c('c', 'd', 'h', 'g','n')), 'q40a', 'q38f')
security_wide_fs <- security_wide[c('school', fs_items)]
```

Analyze Complete Cases
```{r}
# number of no's and yes's from schools
security_wide_fs %>% complete.cases() %>% table()
security_wide %>% dplyr::select(q38c, q38d, q38h, q38g, q38n, q40a, q38f) %>% sapply(function(vec)table(vec, exclude=NULL)) %>% knitr::kable()
# set index for schools based on number of yes responses
sum_index <- security_wide_fs %>% na.omit() %>% dplyr::select(-school) %>% rowSums()
table(sum_index)
```

Format Dataframe
```{r}
security_tall_fs0 <-security_wide_fs %>% na.omit() %>% tidyr::gather("item", "response",-school)
head(security_tall_fs0, 3)
```

# Rasch Model
```{r}
# binomial logistic rasch model
rasch0 <-glm(response~item+school -1,family=binomial, data=security_tall_fs0)
school_coef_names <-paste0("school",levels(security_wide$school))
# display number of yes's and no's from schools
table(school_coef_names %in% names(coef(rasch0)) )
```

Analyze Rasch Indices
```{r}
rasch0_index <-c(schoolid1011=0,coef(rasch0)[-(1:7)])
table(round(rasch0_index))
hist(round(rasch0_index), main="Histogram of Rasch Model 1", xlab="Security Index of Model 1")
rank <- sort(rasch0_index, decreasing=TRUE)
head(rank)
```

# Rasch Model with Different Survey Questions
```{r}
# select questions for analysis
fs_items_c <- c(paste0('q38',c('a', 'e', 'g','h', 'm')), 'q40c', 'q40d')
security_wide_fs_c <- security_wide[c('school', fs_items_c)]
```

Analyze Complete Cases
```{r}
security_wide_fs_c %>% complete.cases() %>% table()
sum_index_c <- security_wide_fs_c%>% na.omit()%>%dplyr::select(-school)%>% rowSums()
table(sum_index_c)
```

Format Dataframe
```{r}
# reformat dataframe
security_tall_fs0_c <-security_wide_fs_c%>% na.omit()%>%tidyr::gather("item", "response",-school)
head(security_tall_fs0_c, 3)
```

Rasch Model
```{r}
rasch0_c <-glm(response~item+school -1,family=binomial, data=security_tall_fs0_c)
head(coef(rasch0_c), 10)
```

Analyze Rasch Indices
```{r}
# set school security indices from Rasch model with selected subset of questions
rasch0_index_c <-c(schoolid1011=0,coef(rasch0_c)[-(1:7)])
table(round(rasch0_index_c))
# display histogram of security indices
hist(round(rasch0_index_c), main="Histogram of Rasch Model 2", xlab="Security Index of Model 2")
```

# Comparison of Models
```{r}
# 629 elements in intersection
common_ids = sort(intersect(names(rasch0_index), names(rasch0_index_c)))
r0_trunc = rasch0_index[common_ids]
r1_trunc = rasch0_index_c[common_ids]
ranks0 = rank(r0_trunc, ties.method = "random")
ranks1 = rank(r1_trunc, ties.method = "random")
```

Plot Comparison of Ranks
```{r}
# plot school rank from model 1 and model 2
# each school is a data point
ggplot(data.frame(cbind(ranks0, ranks1)),aes(ranks0, ranks1)) + 
  geom_point() + ggtitle("Comparison of Model 1 to Model 2") +
  xlab('Model 1 Ranks') + ylab('Model 2 Ranks')
```

Correlation of Model Ranks
```{r}
# 1 indicates correlation
cor(ranks0, ranks1, method ="spearman")
cor(ranks0, ranks1, method ="kendall")
```

The two models are closer to no correlation. The FS model is unreliable but valid.


# Compare Fixed Effects to Random Effects
```{r}
security_tall_fs1 <-security_wide_fs%>%tidyr::gather("item", "response",-school)
rasch01 <-glm(response~item+school -1,family=binomial, data=security_tall_fs1)
rasch01_index <- c(schoolid1011=0,coef(rasch01)[-(1:7)])

rasch1 <- glmer(response~item+(1|school),family=binomial, data=security_tall_fs1)
head(coef(rasch1)$school, 3)
# security indices are now intercept values
rasch1_index <-coef(rasch1)$school[,"(Intercept)"]
sum(security_wide$school != row.names(coef(rasch1)$school))
plot(rasch01_index, rasch1_index, main = "Comparing Variations of Model 1", xlab="Rasch fixed effects model",ylab="Rasch random effects model")

security_tall_fs1_c <-security_wide_fs_c%>%tidyr::gather("item", "response",-school)
rasch01_c <-glm(response~item+school -1,family=binomial, data=security_tall_fs1_c)
rasch01_index_c <- c(schoolid1011=0,coef(rasch01_c)[-(1:7)])

rasch1_c <-glmer(response~item+(1|school),family=binomial, data=security_tall_fs1_c)
head(coef(rasch1_c)$school, 3)
rasch1_index_c <-coef(rasch1_c)$school[,"(Intercept)"]
plot(rasch01_index_c, rasch1_index_c, main = "Comparing Variations of Model 2",xlab="Rasch fixed effects model",ylab="Rasch random effects model")
```

# Bootstrap for random effects personalized model
```{r boot-block, cache=2}
bootreps = 1000
indices = sample(1:nrow(security_tall_fs1_c), size = 3)
ids = as.character(security_tall_fs1_c[, "school"][indices])
# Printing school ids to analyze
ids

rboot = function(statistic, simulator, B)
{
  tboots = replicate(B, statistic(simulator()))
  if(is.null(dim(tboots))) 
  {
    tboots = array(tboots, dim = c(1, B))
  }
  return(tboots)
}

resampler = function()
{
  whichrows = sample(1L:nrow(security_tall_fs1_c), nrow(security_tall_fs1_c), replace = T)
  security_tall_fs1_c[whichrows, ]
}

coefs = function(bootspl)
{
  refitted.mod = update(rasch1_c, data = bootspl)
  schools = coef(refitted.mod)$school
  vals = schools[, "(Intercept)"]
  names(vals) = row.names(schools)
  unlist(vals)
}
boot.stats <- rboot(coefs, resampler, bootreps)

vec1 = numeric(length = bootreps)
vec2 = numeric(length = bootreps)
vec3 = numeric(length = bootreps)
for (i in 1:bootreps)
{
  vec1[i] = unlist(boot.stats[i])[ids[1]]
  vec2[i] = unlist(boot.stats[i])[ids[2]]
  vec3[i] = unlist(boot.stats[i])[ids[3]]
}
vec1 = vec1[!is.na(vec1)]
vec2 = vec2[!is.na(vec2)]
vec3 = vec3[!is.na(vec3)]

#Test for normality and variances
qqnorm(vec1)
hist(vec1, main = paste("Histogram of School" , ids[1]), xlab = "Predicted Index")
# Variance of bootstrapped indices
var(vec1)
# Calculate sample sd of residuals
boot_truth <- coef(rasch1_c)$school[ids[1], "(Intercept)"]
sd_1 <- sd(vec1 - boot_truth)
sd_1

qqnorm(vec2)
hist(vec2, main = paste("Histogram of School" , ids[2]), xlab = "Predicted Index")
# Variance of bootstrapped indices
var(vec2)
# Calculate sample sd of residuals
boot_truth <- coef(rasch1_c)$school[ids[2], "(Intercept)"]
sd_2 <- sd(vec2 - boot_truth)
sd_2

qqnorm(vec3)
hist(vec3, main = paste("Histogram of School" , ids[3]), xlab = "Predicted Index")
# Variance of bootstrapped indices
var(vec3)
# Calculate sample sd of residuals
boot_truth <- coef(rasch1_c)$school[ids[3], "(Intercept)"]
sd_3 <- sd(vec3 - boot_truth)
sd_3
```

```{r}
boxplot(c(sd_1, sd_2, sd_3))
avgsd <- mean(c(sd_1, sd_2, sd_3))
avgsd
# difference between average residual sd and residual sd for each school
c(sd_1, sd_2, sd_3) - avgsd
```

